Algorithm explanation

### 2013-04-18 article/algo.tex
Revised explanation for clarity

Given an acyclic Horn graph (or called a Horn DAG), we define the
relation $v <^+ v'$ by
\[ v <^+ v' \Longleftrightarrow (v,v') \in E \]
By the assumption that $G$ is acyclic, $<^+$ is a well-founded
relation.  We assume without loss of generality that $V$ contains the
least element $v_\bot$ with respect to $<^+$; otherwise we can always
ignore $v_\bot$ with $\varphi(v_\bot) = true$.

For each $v \in V$, we define $\Delta(v) = (a_v x \leq b_v, C_v)$,
where
\begin{itemize}
\item $a_v x \leq b_v$ is a linear inequality that may contain
  coefficient variables
\item $C_v$ is a set of linear inequalities on coefficient variables.
\end{itemize}
Intuitively, the pair $(\mathbf{a} \mathbf{x} \leq b, C)$ denotes the
set of inequalities:

\[ \left\lbrace \mathbf{a} \mathbf{x} \leq b | \exists \neg a,\neg
b. C \right\rbrace \]

(where $\exists \neg a, \neg b.$ is the existential quantification
over coefficient variables other than $a$ and $b$.

$\Delta(v)$ is defined by well-founded induction on $v$ with respect
to the relation $<^+$.

Then,

\begin{align*}
\Delta(v) =
\begin{cases}
\begin{split}
\Bigg(
\sum_i { a_i a_{0,i} x_i } \leq b b_0,
\bigwedge_i \left( - a_i a_{0,i} + \sum_{\substack{(u,v) \in E \\ \Delta(u)=\mathbf{a}_u \mathbf{x} \leq b_u}} a_{u,i} = 0 \right) \wedge \\
\left( \sum_{\substack{(u,v) \in E \\ \Delta(u)=\mathbf{a}_u \mathbf{x} \leq b_u}} b_u - b (b_0+1) < 0 \right) \wedge \\
\bigwedge_{(u,v) \in E} C_u
\Bigg)
& \mbox{if } \varphi(v) = \mathbf{a}_v \mathbf{x} \leq b_v
\end{split}
 \\
\begin{split}
\Bigg(
\mathbf{a} \mathbf{x} \leq b,
\bigwedge_i \left( - a_i + \sum_{\substack{(u,v) \in E \\ \Delta(u)=\mathbf{a}_u \mathbf{x} \leq b_u}} a_{u,i} = 0 \right) \wedge \\
\left( \sum_{\substack{(u,v) \in E \\ \Delta(u)=\mathbf{a}_u \mathbf{x} \leq b_u}} b_u - (b+1) < 0 \right) \wedge \\
\bigwedge_{(u,v) \in E} C_u
\Bigg)
& \mbox{if } \varphi(v) = P(\mathbf{x}).
\end{split}
\end{cases}
\end{align*}

where $\mathbf{a}$ and $b$ are fresh coefficient variables.

If $C_{v_\bot}$ is satisfiable, then any solution $\sigma$ for it
gives a simple solution for $G$, i.e., we should have:

If $\models \sigma(C_{v_\bot})$, then $\rho$ defined by
\[ \rho(v)= \sigma(a_v x <= b_v) \]
is a solution of $G$.

>>>>

Given an acyclic Horn graph (or called a Horn DAG), we define the
relation $v <^+ v'$ by
\[ v <^+ v' \Longleftrightarrow (v,v') \in E \]
By the assumption that $G$ is acyclic, $<^+$ is a well-founded
relation.  We assume without loss of generality that $V$ contains the
least element $v_\bot$ with respect to $<^+$; otherwise we can always
ignore $v_\bot$ with $\varphi(v_\bot) = 0 \leq 0$.

For each $v \in V$, we define $\Delta(v) = (a_v x \leq b_v, C_v)$,
where
\begin{itemize}
\item $a_v x \leq b_v$ is a linear inequality that may contain
  coefficient variables
\item $C_v$ is a set of linear expressions on coefficient variables.
\end{itemize}
Intuitively, the pair $(\mathbf{a} \mathbf{x} \leq b, C)$ denotes the
set of inequalities:

\begin{align*}
\left\lbrace
 \mathbf{a} \mathbf{x} \leq b |
 \exists \left( FV(C)
  \setminus (\mathbf{a} \cup \left\lbrace b \right\rbrace
 \right). C
\right\rbrace
\end{align*}

$\Delta(v)$ is defined by well-founded induction on $v$ with respect
to the relation $<^+$. Then,
$\Delta(v) = \left( \mathbf{a}_v \mathbf{x} \leq b_v, C_v \right)$,
where

\begin{align*}
C_v =
\begin{cases}
\left\lbrace \mathbf{a}_v = \sum_{(u,v) \in E} \mathbf{a}_u \right\rbrace \cup
\left\lbrace b_v >= \sum_{(u,v) \in E} b_u \right\rbrace \cup
\bigcup_{(u,v) \in E} C_u
& \mbox{if } \varphi(v) = P(\mathbf{x}) \vee v = v_\bot \\
\left\lbrace \mathbf{a}_v = \lambda_v \mathbf{a}_0 \right\rbrace \cup
\left\lbrace b_v = \lambda_v b_0 \right\rbrace
& \mbox{if } \varphi(v) = \mathbf{a}_0 \mathbf{x} \leq b_0
\end{cases}
\end{align*}

Given $\varphi(v_\bot) = \mathbf{a}_0 \mathbf{x} \leq b_0$, any
solution $\sigma$ to the constraint $C_\star = C_{v_\bot} \cup {
  \mathbf{a}_0 = \mathbf{a}_{v_\bot} } \cup { b_0 = b_{v_\bot} }$
gives a simple solution for $G$ if $C_\star$ is satisfiable, i.e., we
should have:

\quote
If $\models \sigma(C_\star)$, then $\rho$ defined by
\begin{align*}
 \rho = \left\lbrace
  \left( P, \sigma(\mathbf{a}_v \mathbf{x} \leq b_v) \right) |
  \forall v \in V. \varphi(v) = P(\mathbf{x})
 \right\rbrace
\end{align*}
is a solution of $G$.
\endquote
